---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
setwd("~/rworkspace/reproducibleresearch/peer-assessments1/RepData_PeerAssessment1")
data <- read.csv(unz("activity.zip", "activity.csv"))
```
- Head of the activity dataset
```{r}
head(data)
```

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day
```{r}
result1 <- tapply(data$steps,data$date,sum)
```
- Data for total number of steps taken per day
```{r}
result1
```
2.  Make a histogram of the total number of steps taken each day
```{r}
result1 <- tapply(data$steps,data$date,sum)

hist(result1,xlab="Total daily steps", breaks=25,main="Total number of steps taken per day")
```

3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
total_daily_steps <- as.numeric(tapply(data$steps, data$date, sum))
steps.mean <- mean(total_daily_steps,na.rm= TRUE)
steps.median <- median(total_daily_steps,na.rm=TRUE)
```
- Mean of the total number of steps per day
```{r}
steps.mean
```
- Median of the total number of steps per day
```{r}
steps.median

```

## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
data$interval <- as.factor(as.character(data$interval))
interval.mean <- as.numeric(tapply(data$steps, data$interval, mean, na.rm = TRUE))
intervals <- data.frame(intervals = as.numeric(levels(data$interval)), interval.mean)
intervals <- intervals[order(intervals$intervals), ]

labels <- c("00:00", "05:00", "10:00", "15:00", "20:00")
labels.at <- seq(0, 2000, 500)
plot(intervals$intervals, intervals$interval.mean, type = "l", main = "Average steps 5-minute interval", 
    ylab = "Average steps", xlab = "Time of day", xaxt = "n")
axis(side = 1, at = labels.at, labels = labels)
```


2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
intervals.sorted <- intervals[order(intervals$interval.mean, decreasing = TRUE), 
    ]
head(intervals.sorted)

max.interval <- intervals.sorted$intervals[1[1]]
max.interval
```
- Therefore, the interval between 08:35 and 08:40 minutes has the maximum number of step

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
dim(data[is.na(data$steps), ])[1]
```
- The total number of missing values in the dataset (i.e. the total number of rows with NAs) is 2304.


2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

- The strategy for filling in all of the missing values in the dataset is to change the “NA"s to the mean values for that 5-minute interval.

```{r}
steps <- vector()
for (i in 1:dim(data)[1]) {
    if (is.na(data$steps[i])) {
        steps <- c(steps, intervals$interval.mean[intervals$intervals == data$interval[i]])
    } else {
        steps <- c(steps, data$steps[i])
    }
}


```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activity_without_missing_data <- data.frame(steps = steps, date = data$date, 
    interval = data$interval)
```

- Head of the dataset with the missing data filled in 

```{r}
head(activity_without_missing_data)
```
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
hist(tapply(activity_without_missing_data$steps, activity_without_missing_data$date, 
    sum), xlab = "Total daily steps", breaks = 25, main = "Total of steps taken per day")
```

## Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
activity_without_missing_data$day.type <- c("weekend", "weekday", "weekday", 
    "weekday", "weekday", "weekday", "weekend")[as.POSIXlt(activity_without_missing_data$date)$wday + 
    1]
activity_without_missing_data$day.type <- as.factor(activity_without_missing_data$day.type)

```


- A portion of the new dataset 
```{r}
summary(activity_without_missing_data)
```



2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}

```{r}
weekday <-activity_without_missing_data[activity_without_missing_data$day.type == 
    "weekday", ]
weekend <-activity_without_missing_data[activity_without_missing_data$day.type == 
    "weekend", ]
weekday.means <- as.numeric(tapply(weekday$steps, weekday$interval, mean))
weekend.means <- as.numeric(tapply(weekend$steps, weekend$interval, mean))

intervals_day_type <- data.frame(intervals = as.numeric(levels(data$interval)), 
    weekday.means, weekend.means)
intervals_day_type <- intervals_day_type[order(intervals_day_type$intervals), 
    ]

par <- par(mfrow = c(2, 1))
plot(intervals_day_type$intervals, intervals_day_type$weekday.means, type = "l", 
    col = "red", ylab = "Average steps", xlab = "Time of day", main = "Average steps 5-minute interval at weekday", 
    xaxt = "n")
axis(side = 1, at = labels.at, labels = labels)
plot(intervals_day_type$intervals, intervals_day_type$weekend.means, type = "l", 
    col = "blue", ylab = "Average steps", xlab = "Time of day", main = "Average steps 5-minute interval at weekend", 
    xaxt = "n")
axis(side = 1, at = labels.at, labels = labels)

```